Using VDMS to Search 100M Images: A Piece of Cake

### 2020-02-17 yfcc100m/paper/vldb2020/intro.tex
Minor updates on eval
Varios modification to paper files

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning techniques to
extract information out of the visual data which can subsequently be used
for informed decision making \cite{vdms-nips}.
The insights this information can provide depend on the
application: a retail vendor might be interested in the amount of time
shoppers spend in front of a specific product, while a medical expert might
want to see the effect of a specific treatment on the size of a tumor.

There are many applications that can be studied through the use of large
and public available datasets. But before getting started, there is usually
a large effort devoted to cleaning up the data. For this work, we use
the YFCC100m dataset as an example given that we have use the same dataset
for multiple proof of concepts and applications without our research lab.
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine generated tags for each image, and video summarization.
In all cases, it was needed to select a subset of the data before running
any processing. Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searched
based on tags (relational database, graph database, csv files, etc), and
building the necessary pipelines for querying and retrieving the right data.

Despite this rich and varied usage environment, there has been very little
research on the management of visual data.
Most of the current storage solutions are
an ad-hoc collection of tools combined with custom scripts to tie them
together, unique not only to a specific discipline but often to individual
researchers. For example, consider an ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. This requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database. If the queries require
search over some patient information, then patients have to be associated
with their brain scans. Finally, if the ML pipeline needs images that
are of a size different than the stored ones, there is additional compute
diverted towards preprocessing after the potentially larger images are
fetched. All these steps require investigation of different software
solutions that provide various functionalities that can then be stitched
together with a script for this specific use case.
Moreover, if the pipeline identifies
new metadata to be added for the tumor images, most databases make it
hard to evolve the schema on the fly.
Not only do these ad hoc solutions make replicating experiments
difficult, they do not scale well to real-world applications.
Addressing the storage and retrieval of visual data necessitates a complete
overhaul of the storage architecture,
preferably using emerging breakthroughs in
heterogeneous memory and storage for efficiency.

Data scientist and machine learning developers usually end up building ad-hoc
solutions through a combination of databases and file systems to store metadata
and visual data (images, videos), respectively.

In this paper, we also expand on the Video and Feature Vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyze different functionalities and trade-offs for this type of data,
in combination with metadata filtering. To the best of out knowledge, this
set of functionalities are unique to VDMS and we were not able to find a system
with similar functionality.

@@

\begin{comment}

We present the Visual Data Management System (VDMS)\cite{darkside},
an Open Source project designed to enable efficient access of visual data.
Since visual data often contains
rich metadata (such as objects, locations, and time), VDMS stores this
information in a high performance graph database. Using this metadata, VDMS
can quickly identify which data is relevant to a given query.
Additionally, VDMS uses a custom library to store and retrieve visual data,
which provides an interface for machine friendly formats as well as
traditional formats. These new formats are designed to support applications
that are often interested in specific areas of images or videos,
particularly when the individual object is large.

\end{comment}

>>>>

Visual computing workloads performing analytics on
video or image data, either off-line or streaming,
have become prolific across a wide range of application domains.
This is in part due to the growing ability of machine learning techniques to
extract information out of the visual data which can subsequently be used
for informed decision making and analytics\cite{vdms-nips}.
Visual data can provide important insights for various applications:
a retail vendor might be interested in the amount of time
shoppers spend in front of a specific product, a medical expert might
want to study the effect of a specific treatment on the size of a tumor by
looking at brain scans, and a media company would like to leverage
thousands of hours of produced video material for different uses.

Moreover, there are many applications that can be studied through
the use of large and public available datasets.
But before getting started, there is usually
a large effort devoted to cleaning up the data.
For this work, we use the YFCC100m dataset because of its comprehensive nature,
and mainly because we have used it for multiple proof of concepts
and applications within our research lab at Intel.
Applications include basic image search functionality (through the use
of human-generated tags), advanced image search through the use of
machine generated tags for each image and feature vectors,
and video summarization.
In all cases, it was needed to select a subset of the data before running
any processing. Selecting subsets of data is by itself a time consuming task,
as it involves loading all metadata into a solution that enables searches
based on properties (some of the common tools used for this task are
relational database, graph database, csv files, etc), and
building the necessary pipelines and plug-ins
for querying and retrieving the right data.

Despite the fact that visual data is becoming a primary component in analytics,
there has been very little research on the management of visual data, even
less that resulted in fully functioning and available tools.
Most of the current storage solutions are an ad-hoc collection of
tools combined with custom scripts to tie them together, unique not only to
a specific discipline but often to individual researchers.
For example, consider an ML developer constructing a pipeline
for extracting brain tumor information from existing brain images in a
classic medical imaging use case. This task requires assigning consistent
identifiers for the scans and adding their metadata in
some form of relational or key-value database.
If the queries require search over some patient information,
then patients have to be associated with their brain scans images.
Finally, if the ML pipeline needs images that are of a size different
than the stored ones, there is additional compute diverted
towards preprocessing after the potentially larger images are fetched.
All these steps require investigation of different software
solutions that provide various functionalities that have to be adapted,
re-purposed, and stitched together with scripts for each specific use case.
Moreover, most databases make it hard to evolve the schema on the fly and
require a definition of the representation of the data beforehand, which becomes
a problem when new forms of data becomes available with the advances on ML based
data extraction techniques.
Not only do these ad hoc solutions make replicating experiments
difficult, they do not scale well to real-world applications.
Addressing the storage and retrieval of visual data necessitates a complete
overhaul of the storage architecture, preferably using emerging breakthroughs in
heterogeneous memory and storage for efficiency, and with a focus to recognizing
types of data that becoming the center of analytics (i.e., images and videos).

In this paper, we show how a new approach for visual data management can
a very complex task become simple and productive.
We also show how, buy using a tool that was design for the job,
inefficiencies that emerge from the re-purpose of multiple system disappear
and there are important gains in overall data access performance.
We put a particular emphasis on the scalability and concurrenct aspect of the
performance evaluation given that large datasets are processed in parallel
using distributed computing.

We also expand on the Video and Feature Vector capabilities of
VDMS, which are part of the latest additions to the system.
We analyze different functionalities and trade-offs for this type of data,
in combination with metadata filtering.
To the best of our knowledge, this set of functionalities are unique to
VDMS and we were not able to find a system with similar functionality.

@@

We start by briefly describing VDMS design principles and implementation
in Section \ref{arch}.
We then describe the dataset we used for our experiments and explain
why this is a comprehensive use-case for visual data analytics, as well as
describe our evaluation methodology in Section \ref{eval}.
Finally, we show the results of our evaluation in Section \ref{eval}.

### 2020-02-17 yfcc100m/paper/vldb2020/vdms-vldb2020.tex
Varios modification to paper files

% Include information below and uncomment for camera ready
\vldbTitle{A Sample Proceedings of the VLDB Endowment Paper in LaTeX Format}
\vldbAuthors{Ben Trovato, G. K. M. Tobin, Lars Th{\sf{\o}}rv{$\ddot{\mbox{a}}$}ld, Lawrence P. Leipuner, Sean Fogarty, Charles Palmer, John Smith, Julius P.~Kumquat, and Ahmet Sacan}
\vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
\vldbVolume{12}
\vldbNumber{xxx}
\vldbYear{2019}

@@

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% Authors
\alignauthor
Luis Remis\titlenote{Most of the work was done while at Intel Labs.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{ApertureData Inc.}\\
       \email{luis@aperturedata.io}
\alignauthor
Chaunt\'e W. Lacewell\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
       \affaddr{Intel Labs}\\
       \email{chaunte.w.lacewell@intel.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv\"{a}ld Group, {\texttt{jsmith@affiliation.org}}), Julius P.~Kumquat
%(The \raggedright{Kumquat} Consortium, {\small \texttt{jpkumquat@consortium.net}}), and Ahmet Sacan (Drexel University, {\small \texttt{ahmetdevel@gmail.com}})}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{abstract}
Data scientists spend most of their time dealing with data preparation,
rather than doing what they know best:
build machine learning models and algorithms to solve problems.
The problem grows as the scale of the data that is used for analysis or training
grows. In this paper, we demonstrate how VDMS (Visual Data Management System)
can be used to simplify the data preparation process and we show how we gain
in efficiency just because we are using a tool that was design for the job.
To demonstrate this, we use one of the largest available
public dataset (YFCC100M),
with 99+ million images, 800+ thousand videos plus add-ons that include
machine-generated tags and 4K dimensional feature vectors,
a total of $\sim$13TB of data.
VDMS differs from existing large scale photo
serving, video streaming, and textual big-data management systems due to
its primary focus on supporting machine learning and
data analytics pipelines that use visual data
(images, videos, and feature vectors), treating these as first class citizens.
We demonstrate how VDMS outperforms well-know and widely used tools for
data access by a $\sim$10x factor in certain queries,
while simplifying the process
of data preparation and data access, and providing functionalities non-existent
in alternative options.
\end{abstract}

\input{intro.tex}
\input{design.tex}
\input{eval.tex}
%\input{demo.tex}
\input{conclusion.tex}

@@

\balance
\begin{appendix}
You can use an appendix for optional proofs or details of your evaluation which are not absolutely necessary to the core understanding of your paper.

@@

\end{appendix}

>>>>

% Include information below and uncomment for camera ready
\vldbTitle{Using VDMS to Search 100M Images: A Piece of Cake}
\vldbAuthors{Luis Remis, Chaunt\'e W. Lacewell}
\vldbDOI{https://doi.org/10.14778/xxxxxxx.xxxxxxx}
\vldbVolume{12}
\vldbNumber{xxx}
\vldbYear{2020}

@@

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% Authors
\alignauthor
Luis Remis\titlenote{Luis was a Research Scientist at Intel Labs until late 2019. Most of this work was done while at Intel Labs.}\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{ApertureData Inc.}\\
       \email{luis@aperturedata.io}
\alignauthor
Chaunt\'e W. Lacewell\\
%       \affaddr{Institute for Clarity in Documentation}\\
%       \affaddr{P.O. Box 1212}\\
       \affaddr{Intel Labs}\\
       \email{chaunte.w.lacewell@intel.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv\"{a}ld Group, {\texttt{jsmith@affiliation.org}}), Julius P.~Kumquat
%(The \raggedright{Kumquat} Consortium, {\small \texttt{jpkumquat@consortium.net}}), and Ahmet Sacan (Drexel University, {\small \texttt{ahmetdevel@gmail.com}})}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{abstract}
Data scientists spend most of their time dealing with data preparation,
rather than doing what they know best:
build machine learning models and algorithms to solve problems.
The problem grows as the scale of the data that is used for analysis or training
grows. In this paper, we demonstrate how VDMS (the Visual Data Management System)
can be used to simplify the data preparation process and we show how we gain
in efficiency just because we are using a tool that was design for the job.
To demonstrate this, we use one of the largest available
public dataset (YFCC100M),
with 99+ million images, 800+ thousand videos, plus add-ons that include
machine-generated tags and 4K dimensional feature vectors,
for a total of $\sim$13TB of data.
VDMS differs from existing large scale photo
serving, video streaming, and textual big-data management systems due to
its primary focus on supporting machine learning and
data analytics pipelines that use visual data
(images, videos, and feature vectors), treating these as first class citizens.
We demonstrate how VDMS outperforms well-know and widely used tools for
data access by a $\sim$10x factor on this dataset most of the queries we
used for our internal use-cases, and particularly at scale,
while simplifying the process of data preparation and data access,
and providing functionalities non-existent in alternative options.
\end{abstract}

\input{intro.tex}
\input{design.tex}
\input{eval.tex}
\input{conclusion.tex}

@@

\balance

\begin{comment}
\begin{appendix}
You can use an appendix for optional proofs or details of your evaluation which are not absolutely necessary to the core understanding of your paper.

@@

\end{appendix}
\end{comment}
